### 2017-05-19 docs/thesis/app_fits.tex

\chapter{Convex fitted functions}

Some constraints in this optimization study are convex equations that approximate data or other functions.  
These equations include wind speed vs air density, airfoil drag polar, tail drag polar, BSFC to throttle mapping, and arctan equations.  
Each of these equations was generated using the fitting techniques described by Hoburg et. al.\cite{fitting} 
All fitted equations are either \emph{max-affine} or \emph{softmax-affine} functions,

\begin{align}
    f_{\text{MA}} (\bm{x}) &= \max\limits_{k=1 \dots K} [b_k + \bm{x}_K^T\bm{x}] \\
    f_{\text{SMA}} (\bm{x}) &= \frac{1}{\alpha} \log{\sum_{k=1}^K \exp{(\alpha(b_k + \bm{a}_k^T \bm{x}))}}.
\end{align}

The log-space RMS error,

\begin{equation}
\text{RMS error} \equiv \sqrt{\frac{1}{m} \sum_{i=1}^m (f(\bm{x}_i) - y_i)^2},
\end{equation}

 is given throughout the paper for each equation, where $m$ is the number of data points.

Bounds are imposed inside the optimization to ensure that the optimizier does not find a solution that exceeds the range over which the data was fitted.  In the case of the wing and tail drag polar fitted equations, a post-solve process runs XFOIL and verifies that the optimized drag for a given lift coefficient and Reynolds number is within 5\% of the XFOIL computed drag.