### 2017-05-19 docs/thesis/chap_concl.tex

\chapter{Conclusion}
Using geometric programming the feasibility limits and design space of solar-electric and gas powered aircraft is analyzed is second to a level of detail and speed not previously achieved in initial aircraft sizing and design. 
Based on the assumptions included herein, the gas powered aircraft is more capable of meeting high latitude and station keeping requirements. 
The solar-electric powered aircraft sizing does not converge at latitude requirements above 31 degrees latitude. 
However, solar-electric powered aircraft can meet high endurance and altitude requirements.  
The effect of key design parameters on the size of each architecture was also achieved.  
Using higher energy density batteries can result in significant weight and performance savings for the solar-electric powered aircraft.  
The rapid solve time of geometric programming allows detailed sizing and design analysis of long-endurance aircraft and the capability of solar-electric and gas powered aircraft to meet various requirements.