### 2017-05-22 docs/thesis/cover.tex
change naming for tables; update intro conclusion and result sections; add sp reference

% -*-latex-*-
% 
% For questions, comments, concerns or complaints:
% thesis@mit.edu
% 
%
% $Log: cover.tex,v $
% Revision 1.8  2008/05/13 15:02:15  jdreed
% Degree month is June, not May.  Added note about prevdegrees.
% Arthur Smith's title updated
%
% Revision 1.7  2001/02/08 18:53:16  boojum
% changed some \newpages to \cleardoublepages
%
% Revision 1.6  1999/10/21 14:49:31  boojum
% changed comment referring to documentstyle
%
% Revision 1.5  1999/10/21 14:39:04  boojum
% *** empty log message ***
%
% Revision 1.4  1997/04/18  17:54:10  othomas
% added page numbers on abstract and cover, and made 1 abstract
% page the default rather than 2.  (anne hunter tells me this
% is the new institute standard.)
%
% Revision 1.4  1997/04/18  17:54:10  othomas
% added page numbers on abstract and cover, and made 1 abstract
% page the default rather than 2.  (anne hunter tells me this
% is the new institute standard.)
%
% Revision 1.3  93/05/17  17:06:29  starflt
% Added acknowledgements section (suggested by tompalka)
% 
% Revision 1.2  92/04/22  13:13:13  epeisach
% Fixes for 1991 course 6 requirements
% Phrase "and to grant others the right to do so" has been added to 
% permission clause
% Second copy of abstract is not counted as separate pages so numbering works
% out
% 
% Revision 1.1  92/04/22  13:08:20  epeisach

% NOTE:
% These templates make an effort to conform to the MIT Thesis specifications,
% however the specifications can change.  We recommend that you verify the
% layout of your title page with your thesis advisor and/or the MIT 
% Libraries before printing your final copy.
\title{Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming}

\author{Michael Burton}
% If you wish to list your previous degrees on the cover page, use the 
% previous degrees command:
%       \prevdegrees{A.A., Harvard University (1985)}
% You can use the \\ command to list multiple previous degrees
%       \prevdegrees{B.S., University of California (1978) \\
%                    S.M., Massachusetts Institute of Technology (1981)}
\prevdegrees{B.S., Massachusetts Institute of Technology, 2016}
\department{Department of Aeronautics and Astronautics}

% If the thesis is for two degrees simultaneously, list them both
% separated by \and like this:
% \degree{Doctor of Philosophy \and Master of Science}
\degree{Master of Science in Aeronautics and Astronautics}

% As of the 2007-08 academic year, valid degree months are September, 
% February, or June.  The default is June.
\degreemonth{June}
\degreeyear{2017}
\thesisdate{May 25, 2017}

%% By default, the thesis will be copyrighted to MIT.  If you need to copyright
%% the thesis to yourself, just specify the `vi' documentclass option.  If for
%% some reason you want to exactly specify the copyright notice text, you can
%% use the \copyrightnoticetext command.  
%\copyrightnoticetext{\copyright IBM, 1990.  Do not open till Xmas.}

% If there is more than one supervisor, use the \supervisor command
% once for each.
\supervisor{Warren Hoburg}{Associate Professor of Aeronautics and Astronautics}

% This is the department committee chairman, not the thesis committee
% chairman.  You should replace this with your Department's Committee
% Chairman.
\chairman{Youssef M. Marzouk}{Chairman, Department Committee on Graduate Theses}

% Make the titlepage based on the above information.  If you need
% something special and can't use the standard form, you can specify
% the exact text of the titlepage yourself.  Put it in a titlepage
% environment and leave blank lines where you want vertical space.
% The spaces will be adjusted to fill the entire page.  The dotted
% lines for the signatures are made with the \signature command.
\maketitle

% The abstractpage environment sets up everything on the page except
% the text itself.  The title and other header material are put at the
% top of the page, and the supervisors are listed at the bottom.  A
% new page is begun both before and after.  Of course, an abstract may
% be more than one page itself.  If you need more control over the
% format of the page, you can use the abstract environment, which puts
% the word "Abstract" at the beginning and single spaces its text.

%% You can either \input (*not* \include) your abstract file, or you can put
%% the text of the abstract directly between the \begin{abstractpage} and
%% \end{abstractpage} commands.

% First copy: start a new page, and save the page number.
\cleardoublepage
% Uncomment the next line if you do NOT want a page number on your
% abstract and acknowledgments pages.
% \pagestyle{empty}
\setcounter{savepage}{\thepage}
\begin{abstractpage}
\input{abstract}
\end{abstractpage}

% Additional copy: start a new page, and reset the page number.  This way,
% the second copy of the abstract is not counted as separate pages.
% Uncomment the next 6 lines if you need two copies of the abstract
% page.
% \setcounter{page}{\thesavepage}
% \begin{abstractpage}
% \input{abstract}
% \end{abstractpage}

\cleardoublepage

\section*{Acknowledgments}

The success I achieved in completing this thesis was only possible because of the incredible support from family, friends, mentors, and MIT faculty.
I was able to complete this thesis a full year ealier than I expected and I must thank the many people who made that possible.

I first need to thank my family beginning with my beautiful wife, Jasmine.  
Her support and encouragement have been unwavering.  
When I was up late struggling with a problem, she never believed I wouldn't figure it out.  
She has always been enthusiastic and supportive of my successes and given sound advice and encouragment when facing difficult problems or situations. 
I could not have done this without her.

My parents and sibilings were also constant sources of support and encouragement. 
My parents are the wisest people I know and I am immensely grateful for their advice.  
They have always pushed me to be the best me and I would not have ever made it to MIT without them.

I want to greatly thank my advisor, Warren Hoburg (Woody), for his guidance, friendship, and belief in my abilities.  
Woody is a demanding advisor, who pushed me to learn, grow, and develop at a pace and intensity that would not have been possible otherwise. 
While rigorous in his pursuit of high quality work, Woody also gave me the flexibility to explore my own interests and taught me to think critically for myself.
He is truly a brillant professor and I was fortunate that he selected me to join his lab this year.

I would not have made it through the rigors and stress of MIT without my best friend and officemate Martin York. 
I could not have shared an office with a more brillant, passionate and fun person. 
Our common interests, beliefs, and mentality gave me freedom to express my views, opinions, challenges and successes without fear of judgement or offense.

I also need to thank the project managers, professors, and sponsors of the Jungle Hawk Owl, John Hansmen, Navid Yazdani, Mark Drela, Warren Hoburg and Tony Tao. 
The original problem developed by this team of professionals, of which architecture would was best for the Jungle Hawk Owl was the basis for my thesis work.
They all acted as mentors to me, teaching me techincally, guiding me in the industry and giving me opportunites to present my work. 
It was a privilege to work with and for these people to build the Jungle Hawk Owl.

Finally, I want to thank the many fellow students I worked with over the past year and a half building and designing the Jungle Hawk Owl.  
We went through an extremely intense and rapid design and build cycle. 
The late nights, long, arduous layups, and successful test flights have forged a camaraderie between us that I will not forget.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% -*-latex-*-